---
title: "Predicting with confidence"
author: "Katarzyna Kryńska, Jerzy Bednarski, Michał Szałanski"
output: html_document
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(fig.width = 16)
knitr::opts_chunk$set(fig.height = 9)
```

```{r to_do, eval = F}
To-do:
  add theoretical content
  add analysis with missing data (as suggested)
```

# Problem definition 
What is the problem (from the presentation)


# Theoretical background and literature review
About our solution (generally, like the presentation) + the paper

# Implementation in R

### Libraries
To start with, we need to load the required libraries. For random forest we'll be using the ranger package. Then we'll use plotting2 to plot the results, with the fantastic theme provided by the director of data science at Airbnb. 
```{r libraries}
library(ranger)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(ggtech) # For the airbnb theme
library(extrafont) # To add the font for Airbnb theme

theme_set(theme_airbnb_fancy())
pink = "#FF5A5F"
orange = "#FFB400"
blueGreen = "#007A87"
flesh = "#FFAA91"
purple = "#7B0051"
options(scipen=999) #avoiding e10 notation
```

### Experiment #1 - uneven groups
To start, we wanted to see the effect of a different number of observations within each group. If the model were to behave as expected, the group with fewer observations should have a wider margin of confidence.   

To test this, we prepared an artificial dataset. We try to predict the default probability as a factor of the type of employment. The two options are: salary and self employed. The analysis is based on historical data and such we have an imbalanced set: 10 000 people on salary, and only 500 self employed.   

The results are promising - the less populated group does indeed have a much higher margin of error. It also seems to follow the general limit theorem - the number of observations in the salaried group is 20 times bigger, but the variance is only about 4.5 times smaller. 

```{r uneven_groups, cache = TRUE}
# Create the articilial dataset with uneven groups
didDefault <- c(c(rep(1, 165) , rep(0, 10000 - 165)),
                c(rep(1, 57), rep(0, 500 - 57)))

self_employed <- c(rep(0, 10000), rep(1, 500))
data <- tibble(self_employed=factor(self_employed), didDefault = factor(didDefault))

# Run the Random Forest
rf <- ranger("didDefault ~ .", data=data, probability = T, keep.inbag=TRUE)

# Predict the results, including the sd
res2 <- predict(rf, data, type ='se')
res2 <- tibble(pred_prob = res2$predictions[,2], se = res2$se[,2])
res2 <- cbind(data ,res2)

# Group the results and add the confidence intervals
# Then plot 
res2 %>% group_by(self_employed) %>% 
  summarise(n_def = sum(didDefault == 1),
    n = n(),
    pd = mean(pred_prob),
            se = sqrt(mean(se^2)),
            pd_low = pd - 1.96 * se,
            pd_up = pd + 1.96 * se) %>%
  ggplot(aes(x = factor(self_employed), y = pd)) +
  geom_point(color = blueGreen, size = 7) +
  geom_errorbar(aes(ymin = pd_low, ymax = pd_up), size = 3, width = 0.1, color = blueGreen, alpha = 0.5) + 
  labs(
    title = 'Probability of default - type of employment',
    subtitle = 'Salary: 10 000, Self employed: 500',
    x = '',
    y = 'Probability of default'
  ) + 
  scale_x_discrete(labels = c('Salary (n = 10 000)', 'Self employed (n = 500)')) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.16)) 
```

### Experiment #2 - even groups

In the previews experiment, the default probability was very different for the two groups. To exclude the possibility of this influencing the variance, we prepared a second experiment. This time the groups are more or less even, but the default rate is different.   

The data is also artificial - we are simulating the influence of marital status on the pd. We have 10 000 observations of married people and 9 000 singles.  

And the results are also promising - the groups seem to have similar margins of confidence, despite having very different probabilities of default. 

```{r even_groups, cache = TRUE}
# Like above, but with even groups
didDefault <- c(c(rep(1, 165) , rep(0, 10000 - 165)),
                c(rep(1, 320), rep(0, 9000 - 320)))
self_employed <- c(rep(0, 10000), rep(1, 9000))
data <- tibble(self_employed=factor(self_employed), didDefault = factor(didDefault))

rf <- ranger("didDefault ~ .", data=data, probability = T, keep.inbag=TRUE)

res2 <- predict(rf, data, type ='se')
res2 <- tibble(pred_prob = res2$predictions[,2], se = res2$se[,2])
res2 <- cbind(data ,res2)

res2 %>% group_by(self_employed) %>% 
  summarise(n_def = sum(didDefault == 1),
            n = n(),
            pd = mean(pred_prob),
            se = sqrt(mean(se^2)),
            pd_low = pd - 1.96 * se,
            pd_up = pd + 1.96 * se) %>%
  ggplot(aes(x = factor(self_employed), y = pd)) +
  geom_point(color = blueGreen, size = 7) +
  geom_errorbar(aes(ymin = pd_low, ymax = pd_up), size = 3, width = 0.1, color = blueGreen, alpha = 0.5) + 
  labs(
    title = 'Probability of default - marital status',
    subtitle = 'Married: 10 000, Single: 9 000',
    x = '',
    y = 'Probability of default'
  ) + 
  scale_x_discrete(labels = c('Married (n = 10 000)', 'Single (n = 9 000)')) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.042)) 
```

### Experiment #3 - real data
To push the experiments further, we searched for a real dataset. We found a promising database from Taiwan, consisting of 30 000 observations and 23 variables.  

After importing the data, cleaning it and removing outliers, we imported it into the model. This time, we are taking a slightly different approach. We are running a 3-fold cross validation on this data to get the OOS probability and variance. 

For the plot below, we are using a new metric - "Model confidence". Since this is a binary classification problem, we calculate the model confidence as (1 - standard error) * 100%. Thous 100% means that the model is absolutely confident in it's prediction and lower means that it starts to loose confidence. It has to be noted, that we calculate standard error and PD for each observation individually.   

The results are very interesting - it seems that the model is very confident when it predicts PD near 0% and 100%, but it starts to loose confidence for some observations when the predicted probability is closer to 50%. 

```{r real_data, eval = TRUE, cache = T}
# Read processed data
# The code for processing can be found on GitHub, not included here for brevity
# Includes outlier removal and data cleaning
data <- read_csv("data/processed/credit_default_clean.csv")

# Remove the id column
data <- data[,-1]

target <- 'didDefault'
data[,target] <- factor(data[[target]])

# This function runs k fold cross validation and append following columns to original data:
# id - row number from data
# pred_prob - OOS predicted probability of default
# se - estimated OOS standard error
rf_cv <- function(data, n_folds=3){
  form <- formula("didDefault ~ .")
  
  folds <- sample(rep(c(1:n_folds),ceiling(nrow(data)/n_folds)), size = nrow(data))
  results <- list()
  
  for(i in 1:n_folds){
    # Build the model on training data
    rf_fit <- ranger(formula=form, 
                     data=na.omit(data[folds != i ,]), 
                     probability=TRUE,
                     keep.inbag=TRUE) 
    
    # Generate predictions and standard errors for test set
    pred <- predict(rf_fit, data=na.omit(data[folds == i,]), type = "se")
    
    results[i] <- list(as_tibble(list(id = which(folds==i), 
                                      pred_prob = pred$predictions[,2], 
                                      se = pred$se[,2])))
    
  }
  results <- do.call(rbind, results)
  results <- results %>% arrange(id)
  results <- cbind(data, results)
  return(results)
}

# Run the CV
res <- rf_cv(data, 3)

res$trueResult <- as.integer(res$didDefault) - 1

# Plot the results
ggplot(res) +
  geom_jitter(aes(x = pred_prob, y = 1 - se, color = 1 - se), alpha = 0.4) +
  scale_color_gradient(high = orange, low = purple) +
  labs(title = 'Model confidence vs predicted probability of default',
       subtitle = 'Mean probability of default: 26.7%',
       x = 'Probability of default',
       y = 'Model confidence') +
  theme(legend.position = 'none') +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1))
```

### Experiment #4 - missing data

```{r eval = F}
HERE add the missing data experiment (remember to remove the eval = F)
```

# Potential uses and summary

# Sources 
Add the main one + what the dude had in his presentation

### Confidence intervals for Random Forest
Wager, S., Hastie T., & Efron, B. (2014). Confidence Intervals for Random Forests: The Jackknife and the Infinitesimal Jackknife. J Mach Learn Res 15:1625-1651

### Dataset
Yeh, I. C., & Lien, C. H. (2009). The comparisons of data mining techniques for the predictive accuracy of probability of default of credit card clients. Expert Systems with Applications, 36(2), 2473-2480.
https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients
